Individual foraging specialization and success change across experience in a virtual predator-prey system

### 2025-02-21 manuscript/manuscript.Rmd

In this study, we investigate how repeated experience shapes predator foraging specialization and success using data from players in *Dead by Daylight*. In this videogame, four prey players forage for resources while avoiding predation by a fifth player. The predator population is composed of individuals that either move at slow speeds and ambush or hunt at high speeds (i.e. mean movement speed describes an ambush-cursorial continuum of hunting tactics) [@@fraserfranco.etal2022]. We also observed that predators acquire expertise through extensive practice, which is driven by the movement of the prey [@fraserfranco.etal2024]. Indeed, the prey can increase their chances of survival by moving faster in the virtual environment [@fraserfranco.etal2022;@fraserfranco.etal2024].

@@

@@

knitr::include_graphics(
  here::here("outputs", "04_outputs_figures", "figure4.png")
)

@@

The predator's mean speed and IIV were only weakly negatively correlated with hunting success across experience (Figure 3), suggesting that slower/flexible and faster/specialist predators were overall both equally successful. There was however a slight increase in the correlation between the predator's IIV in speed and success (Figure 3C), suggesting that flexible foragers at the advanced stage were marginally less successful (Appendix 1: Figure S2). We also found strong evidence that individuals who encountered groups of prey with faster speeds were less successful (Figure 3A-B). Predators also had greater success across the study when encountering more variable groups of prey (Figure 4A-B). Both relationships remained structurally stable across experience (Figure 3C).

>>>>

In this study, we investigate how repeated experience shapes predator foraging specialization and success using data from players in *Dead by Daylight*. In this videogame, four prey players forage for resources while avoiding predation by a fifth player. The predator population is composed of individuals that either move at slow speeds and ambush or hunt at high speeds (i.e. mean movement speed describes an ambush-cursorial continuum of hunting tactics) [@fraserfranco.etal2022]. We also observed that predators acquire expertise through extensive practice, which is driven by the movement of the prey [@fraserfranco.etal2024]. Indeed, the prey can increase their chances of survival by moving faster in the virtual environment [@fraserfranco.etal2022;@fraserfranco.etal2024].

@@

knitr::include_graphics(
  here::here("outputs", "04_outputs_figures", "figure1.png")
)

@@

@@

The predator's mean speed and IIV were only weakly negatively correlated with hunting success across experience (Figure 3), suggesting that slower/flexible and faster/specialist predators were overall both equally successful. There was however a slight increase in the correlation between the predator's IIV in speed and success (Figure 3C), suggesting that flexible foragers at the advanced stage were marginally less successful (Appendix 1: Figure S2). We also found strong evidence that individuals who encountered groups of prey with faster speeds were less successful (Figure 3A-B). Predators also had greater success across the study when encountering more variable groups of prey (Figure 3A-B). Both relationships remained structurally stable across experience (Figure 3C).
